Mssel_analysis

### 2019-06-10 analysis/mssel_analysis.Rmd
Talk Prep Figures

## Load packages and import data
```{r setup, message=F}
knitr::opts_chunk$set(echo = TRUE)
library(assertthat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(PopGenome)
library(stringr)
library(data.table)
library(magrittr)
library(wrapr)
library(foreach)

ms2 = read.table("~/Documents/Research/PloidySim/data/mssel_out_2-24-19_fin.txt", head = T)

#Group windows into larger sets
ms2 %<>% mutate(ints = as.numeric(as.character(cut(bp.start, breaks = 50, labels = seq(1,50)))))

@@

# Example

@@

# Param set to create trajectories for
params = data.frame("s" = rep(0.1,9), "dom" = c(rep(0.5,3), rep(0,3), rep(1,3)), "ploidy" = rep(c(2,4,8),3), "start" = rep(0.05, 9))

@@

#plot
ggplot(dat1, aes(x=gen, y=freq, linetype=as.factor(dom), color=as.factor(ploidy))) + geom_line() + scale_color_discrete(name="Ploidy") + scale_linetype_discrete(name="Dominance Scalar", labels=c("0.1", "0.5", "0.9")) + xlim(0,300) + theme_bw() + xlab("Generation") + ylab("Allele Frequency")
```

@@

## Summary of depth and breadth of peaks {.tabset}
Controlling for number of chromosomes in the population (i.e. Simulation Ploidy), higher ploidies have a shallower dip, due to the slowed rate of fixation of the sweeping allele. 
Similarly, the dips in diversity are narrower in higher ploidies for same reason as before.

### Depth
```{r, echo=T,message=F}
MS2.a = ms2 %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep) %>% summarize(depth = (mean(Pi.2,na.rm=T) - min(Pi.1)) / mean(Pi.2,na.rm=T))

MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=-(log(depth)), fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot() + ylab(expression(paste("Dip Depth = min(", pi[1], ") / ",pi[2],sep = ""))) + facet_grid(~sampGen) + scale_y_log10() + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()
```

### Breadth
```{r, echo=T,message=F}
MS2.b = ms2 %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, ints) %>% summarize(lessWind = ifelse(mean(Pi.1, na.rm=T) < (mean(Pi.2, na.rm=T) / 2),1,0))

MS2.c = MS2.b %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep) %>% summarize(breadth = sum(lessWind) * 20000)

@@

@@

```{r, echo=T,message=F}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & N == 10000 & fuseGen == 1 & sampGen ==1 & mu==1e-08 & sim_ploidy == 8) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Trajectory\nPloidy") + facet_grid(vars(s, recomb), vars(dom),labeller = label_both, scales = "free_y") + theme_bw()
```

@@

ms2 %>% filter(dom == "0.5" & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb == 1e-08) %>% ggplot(., aes(y=Tajima.D_1, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Tajima's D") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(traj_ploidy),labeller = label_both) + theme_bw()
```

@@

Neutral divergence in allele frequencies increases Fst more quickly in lower ploidies, thus erasing 'outlier' signal due to sweep
```{r, echo=T,message=F}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & sim_ploidy == traj_ploidy & recomb == 1e-08) %>% ggplot(., aes(y=fst, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Fst") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()
```

# Analysis of rehh calculations

This R package calculates EHH (Extended Haplotype Homozygosity) and related metrics, which can be used to detect a selective sweep.  I'm showing results for the XP-EHH, the cross-populatin EHH.

[Link to Article](https://www.nature.com/articles/nature06250)

### Import Data
```{r}
traj.ehh = read.csv("~/Documents/Research/PloidySim/rehh/rehh_calcs_sp2_se0.01.csv")
fact.ehh = read.csv("~/Documents/Research/PloidySim/rehh/rehh_calcs_do0.5_fG1_DS.csv")
```

## Trajectory effect

Faster sweeps in the lower ploidies results in stronger signals of selection.
```{r, echo=T,message=F}
traj.ehh %>% filter(fuseGen=="fG1" & sampGen=="sG1" & recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION / 100000,y=xpehh.p,color=traj_ploidy)) + geom_smooth(method="loess", span=0.1, se = F) + scale_color_discrete(name="Trajectory\nPloidy", breaks=c("tp2","tp4","tp8"), labels=c("2","4","8")) + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()
```

## Factor-of-two effect

Trajectory ploidy is constant within each panel, while simulation ploidy is varied.  Higher diversity in higher ploidies produces stronger signal due to greater deviation from baseline genomic diversity.

Something strange here though...why is p value always highest when tp=sp?
Maybe look at density of NA along chromosomes?
```{r, echo=T,message=F}
fact.ehh %>% filter(s == "se0.01" & fuseGen=="fG1" & recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION / 100000, y=xpehh.p, color=sim_ploidy)) + geom_smooth(method="loess", span=0.1, se = F) + scale_color_discrete(name="Simulation\nPloidy", breaks=c("sp2","sp4","sp8"), labels=c("2","4","8"))  + facet_grid(cols = vars(traj_ploidy), rows = vars(sampGen))  + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()
```

## Sample generation effect

Does signal decay faster in higher or lower ploidies?
Answer seems to be that decay is faster in lower ploidies.  Drift is faster in lower ploidies, which is primary mechanism that erases signal of selection.  
```{r}
fact.ehh %>% filter(s == "se0.01" & fuseGen=="fG1" & dom=="do0.5" & N=="NN10000" & recomb=="re1e-08") %>% ggplot(.,aes(x=POSITION,y=xpehh.p,color=sim_ploidy)) + geom_smooth() + facet_grid(rows = vars(traj_ploidy), cols = vars(sampGen))
```

# Less interesting analyses

### Effect of sample generation
Sample generation is the number of generations following a sweep at which sampling occurs
```{r} 
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb == 5e-08 & sim_ploidy == traj_ploidy) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + scale_y_continuous(breaks=c(0.05,0.15)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Simulation\nPloidy") + facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()
```

### Effect of population size
Nothing really interesting here...Smaller population size seems noisier as expected
```{r}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & recomb=="1e-08" & fuseGen == 1 & sampGen ==1 & mu==1e-08 & sim_ploidy == traj_ploidy) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Ploidy") + facet_grid(vars(s, N), vars(dom),labeller = label_both, scales = "free_y") + theme_bw()
```

### Other plots
```{r}
fact.ehh %>% filter(traj_ploidy == "tp4" & fuseGen=="fG1" & dom=="do0.5" & N=="NN10000" & recomb=="re1e-08" & sampGen=="sG1") %>% ggplot(.,aes(x=cut_interval(x=POSITION, n=50),y=xpehh.p,fill=sim_ploidy)) + geom_boxplot() + facet_grid(cols = vars(sampGen))

# # Look for proportional difference in breadth/depth within a simul
MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & traj_ploidy == 4 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=depth, fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(notch=T) + ylab(expression(paste("Dip Breadth = Length of which ", pi[1], " < (",pi[2], " / 2)",sep = "")))+ scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()

fact.ehh %>% filter(sim_ploidy=="sp8" & fuseGen=="fG1" & sampGen=="sG1" & recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION / 100000,y=xpehh.p,color=traj_ploidy)) + geom_smooth(method="loess", span=0.1, se = F) + scale_color_discrete(name="Trajectory\nPloidy", breaks=c("tp2","tp4","tp8"), labels=c("2","4","8")) + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()
```

>>>>

## Load packages and import data
```{r setup, message=F}
knitr::opts_chunk$set(echo = TRUE)
library(assertthat)
library(ggplot2)
library(dplyr)
library(tidyr)
library(PopGenome)
library(stringr)
library(data.table)
library(magrittr)
library(wrapr)
library(foreach)
library(purrr)

ms = read.table("~/Documents/Research/PloidySim/data/mssel_out_2-24-19_fin.txt", head = T)

#Group windows into larger sets
ms %<>% mutate(driftGen = 0)

ms2 = read.table("~/Documents/Research/PloidySim/data/mssel_out_highRe_drift_4-30-19.txt", head = T)

#NEED TO ADD this data
ms3 = read.table("~/Documents/Research/PloidySim/data/mssel_out_5-10-19.txt", head = T)
ms3 %<>% filter(sim_ploidy!=1)

ms = rbind(ms,ms2)

ms %<>% mutate(ints = as.numeric(as.character(cut(bp.start, breaks = 50, labels = seq(1,50)))))
ms3 %<>% mutate(ints = as.numeric(as.character(cut(bp.start, breaks = 50, labels = seq(1,50)))))

ms = rbind(ms, ms3)

@@

mDoms %>% filter(variable %in% c("0.1", "0.5", "0.9")) %>% ggplot(., aes(x = i.k, y = value, linetype = variable)) + geom_line() + scale_color_discrete(name = "Dominance", breaks = c("0.1", "0.5", "0.9"), labels = c("Partial\nRecessivity\n", "Additive\n", "Partial\nDominance")) + theme_bw() + xlab("Individual allele frequency of beneficial allele") + ylab("Dominance Coefficient")

mDoms$variable = factor(mDoms$variable, levels(mDoms$variable)[c(rev(seq(1,11)))])
ALL$Call=factor(ALL$Call,levels(ALL$Call)[c(2,1,3)])

#Color
mDoms %>% filter(variable %in% c("0.1", "0.5", "0.9")) %>% ggplot(., aes(x = i.k, y = value, color = variable)) + geom_line(size = 1.2) + scale_color_manual(name = "Dominance", breaks = c("0.9", "0.5", "0.1"), labels = c("Partial\nDominance", "Additive", "Partial\nRecessivity\n"), values = c("#E6A0C4", "#C6CDF7", "#D8A499")) + theme_bw() + xlab("Individual allele frequency of beneficial allele") + ylab("Dominance Coefficient") + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))

#Linetype
mDoms %>% filter(variable %in% c("0.1", "0.5", "0.9")) %>% ggplot(., aes(x = i.k, y = value, linetype = variable)) + geom_line() + scale_linetype_manual(name = "Dominance", breaks = c("0.1", "0.5", "0.9"), labels = c("Partial\nRecessivity\n", "Additive\n", "Partial\nDominance"), values = c("dashed", "solid","dotdash")) + theme_bw() + xlab("Individual allele frequency of beneficial allele") + ylab("Dominance Coefficient") + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))
# Example

@@

# Param set to create trajectories for
params = data.frame("s" = rep(0.1,9), "dom" = c(rep(0.5,3), rep(0.1,3), rep(0.9,3)), "ploidy" = rep(c(2,4,8),3), "start" = rep(0.05, 9))

@@

dat1$dom = factor(dat1$dom, levels(dat1$dom)[c(3,2,1)])
#plot
ggplot(dat1, aes(x=gen, y=freq, linetype=as.factor(dom), color=as.factor(ploidy))) + geom_line() + scale_color_discrete(name="Ploidy") + scale_linetype_discrete(name="Dominance Scalar", labels=c("0.1", "0.5", "0.9")) + xlim(0,300) + theme_bw() + xlab("Generation") + ylab("Allele Frequency")

#ALL
dat1 %>% filter(ploidy %in% c(2, 4, 8) & dom %in% c(0.1, 0.5, 0.9)) %>% 
ggplot(., aes(x=gen, y=freq, linetype=as.factor(dom), color=as.factor(ploidy))) + geom_line() + scale_color_manual(name = "Ploidy", values = c("#F8766D", "#00BA38", "#619CFF")) + scale_linetype_manual(name = "Dominance", labels=c("Partial\nRecessivity\n", "Additivity", "Partial\nDominance")) + xlim(0,300) + theme_bw() + xlab("Generation") + ylab("Allele Frequency") + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))

@@

## Summary of depth and breadth of dips in diversity {.tabset}
Controlling for number of chromosomes in the population (i.e. Simulation Ploidy), higher ploidies have a shallower dip, due to the slowed rate of fixation of the sweeping allele. 
Similarly, the dips in diversity are narrower in higher ploidies for same reason as before.

### Depth
```{r, echo=T,message=F}
MS2.a = ms %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen) %>% summarize(depth = (mean(Pi.2,na.rm=T) - (mean(Pi.2,na.rm=T) - min(Pi.1))) / mean(Pi.2,na.rm=T), min = min(Pi.1, na.rm=T))

MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=-(log(depth)), fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot() + ylab(expression(paste("Dip Depth = [", bar(pi[2])," - (", bar(pi[2]), " - min(", pi[1], ")] / ", bar(pi[2]),sep = ""))) + facet_grid(rows= vars(sampGen), cols = vars(driftGen)) + scale_y_log10() + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()

MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=-(log(min)), fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot() + ylab(expression(paste("Dip Depth = [", bar(pi[2])," - (", bar(pi[2]), " - min(", pi[1], ")] / ", bar(pi[2]),sep = ""))) + facet_grid(rows= vars(sampGen), cols = vars(driftGen)) + scale_y_log10() + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()
```

### Breadth
```{r, echo=T,message=F}
MS2.b = ms %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, driftGen, rep, ints) %>% summarize(lessWind = ifelse(mean(Pi.1, na.rm=T) < (mean(Pi.2, na.rm=T) / 2),1,0))

MS2.c = MS2.b %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, driftGen, rep) %>% summarize(breadth = sum(lessWind) * 20000)

@@

### Drift
What happens if we let the beneficial allele drift neutrally for some number of generations prior to the start of selection

The allele is introduced at frequency 1/(C * N), where C is ploidy.  It then drifts for a specified number of generations as if it is in a population of size (C * N).  After this generation, selection starts.  So, a trajectory ploidy of 2 means that drift and selection are occuring as if this was a diploid population.

Since drift happens more quickly in lower ploidies, the alleles that are not lost after X generations tend to be at higher frequency in lower ploidies.  As selection takes off, these higher starting frequencies result in higher diversity following the sweep.  Although the faster selection in lower ploidies promotes stronger signals of sweeps, the higher starting frequencies due to drift effectively erases or reverses the effect of ploidy on selection.
```{r}
sf = read.table("~/Documents/Research/PloidySim/data/drift_start_freqs.txt", header = T)

sf %>% mutate("Drift Generations" = ends, "Pop. Size" = N) %>% ggplot(., aes(fill = as.factor(ploidy), x = start_freq, y = ..scaled..)) + geom_density(alpha=0.5) + facet_grid(rows=vars(`Drift Generations`), cols = vars(`Pop. Size`), labeller = label_both, scales = "free_y") + ylab("") + xlab("Frequency") + ggtitle("Allele frequency distribution of new mutations \nsurviving after drifting for X generations") + scale_fill_discrete(name = "Ploidy") + theme_bw()

MS3.d = ms3 %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen) %>% summarize(depth = (mean(Pi.2,na.rm=T) - (mean(Pi.2,na.rm=T) - min(Pi.1))) / mean(Pi.2,na.rm=T))

MS3.d %>% filter(sampGen == 1 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="5e-08" & dom == 0.5) %>% ggplot(., aes(y=-(log(depth)), fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot() + ylab(expression(paste("Dip Depth = [", bar(pi[2])," - (", bar(pi[2]), " - min(", pi[1], ")] / ", bar(pi[2]),sep = ""))) + facet_grid(cols = vars(driftGen), rows = vars(s), labeller = label_both) + scale_y_log10() + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()

MS3.b = ms3 %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, rep, driftGen, ints) %>% summarize(lessWind = ifelse(mean(Pi.1, na.rm=T) < (mean(Pi.2, na.rm=T) / 2),1,0))

MS3.c = MS3.b %>% group_by(s, dom, recomb, N, mu, sim_ploidy, traj_ploidy, sampGen, fuseGen, driftGen, rep) %>% summarize(breadth = sum(lessWind) * 20000)

MS3.c %>% filter(sampGen == 1 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb=="1e-08" & dom == 0.5 & sampGen == 1) %>% ggplot(., aes(y=breadth/1000, fill = as.factor(traj_ploidy), x = as.factor(sim_ploidy))) + geom_boxplot() + facet_grid(cols = vars(driftGen), rows = vars(s), labeller = label_both, scales = "free_y") + ylab(expression(paste("Dip Breadth = Length (kb) of which ", pi[1], " < (",pi[2], " / 2)",sep = ""))) + scale_fill_discrete(name = "Trajectory\nPloidy") + xlab("Simulation Ploidy") + theme_bw()
```

@@

```{r, echo=T,message=F}
ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & N == 10000 & fuseGen == 1 & sampGen ==1 & mu==1e-08 & sim_ploidy == 4) %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Trajectory\nPloidy") + facet_grid(vars(s, recomb), vars(dom),labeller = label_both, scales = "free_y") + theme_bw()
```

@@

ms2 %>% filter(dom == "0.5" & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb == 5e-08) %>% ggplot(., aes(y=Tajima.D_1, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Tajima's D") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(traj_ploidy),labeller = label_both) + theme_bw()
```

@@

Neutral divergence in allele frequencies increases Fst more quickly in lower ploidies, thus erasing 'outlier' signal due to sweep
```{r, echo=T,message=F}
ms %>% filter(dom %in% c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & sim_ploidy == traj_ploidy & recomb == 5e-08 & driftGen == 0) %>% ggplot(., aes(y=fst, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Fst") + scale_color_discrete(name="Ploidy") + facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()
```

---
# Evolution figures
```{r}
#Dominance curves
mDoms %>% filter(variable %in% c("0.1", "0.5", "0.9")) %>% ggplot(., aes(x = i.k, y = value, linetype = variable)) + geom_line() + scale_linetype_manual(name = "Dominance", breaks = c("0.1", "0.5", "0.9"), labels = c("Partial\nRecessivity\n", "Additive\n", "Partial\nDominance"), values = c("dashed", "solid","dotdash")) + theme_bw() + xlab("Individual allele frequency of beneficial allele") + ylab("Dominance Coefficient") + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18))

#Basic chromosome diversity plot
ms %>% filter(traj_ploidy==sim_ploidy & sampGen == "1" & s == 0.01 & dom == "0.5") %>% ggplot(., aes(y=Pi.1 * 100, color = as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5, span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") + ylab("Diversity (%)") + scale_color_discrete(name="Ploidy") + theme_bw() + theme(axis.text.y=element_text(size=16), axis.title.y=element_text(size=18),axis.text.x=element_text(size=16),axis.title.x=element_text(size=18),legend.text = element_text(size=16), legend.title = element_text(size=18)) + ylim(0, 0.2)
```

 # Analysis of rehh calculations    This R package calculates EHH
(Extended Haplotype Homozygosity) and related metrics, which can be used to
detect a selective sweep.  I'm showing results for the XP-EHH, the
cross-populatin EHH.   [Link to
Article](https://www.nature.com/articles/nature06250)   ### Import Data
```{r} 
traj.ehh = read.csv("~/Documents/Research/PloidySim/rehh/rehh_calcs_sp2_se0.01.csv") 
fact.ehh = read.csv("~/Documents/Research/PloidySim/rehh/rehh_calcs_do0.5_fG1_DS_fmted.csv") 
```

## Trajectory effect   Faster sweeps in the lower ploidies results in stronger signals of selection.   
```{r, echo=T,message=F}   
traj.ehh %>% filter(fuseGen=="fG1" & sampGen=="sG1" &
recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION /
100000,y=xpehh.p,color=traj_ploidy)) + geom_smooth(method="loess", span=0.1, se
= F) + scale_color_discrete(name="Trajectory\nPloidy",
breaks=c("tp2","tp4","tp8"), labels=c("2","4","8")) + xlab("Position (100 kb)")
+ ylab(expression("-log"[10]*"(p-value)")) + theme_bw()
```

# Factor-of-two effect   Trajectory ploidy is constant within each panel, while simulation ploidy is varied.  Higher diversity in higher ploidies produces stronger signal due to greater deviation from baseline genomic diversity. Something strange here though...why is p value always highest when tp=sp? Maybe look at density of NA along chromosomes?

```{r, echo=T,message=F}   
fact.ehh %>% filter(s == "se0.01" & fuseGen=="fG1" &
recomb == "re1e-08" & dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION /
100000, y=xpehh.p, color=sim_ploidy)) + geom_smooth(method="loess", span=0.1, se
= F) + scale_color_discrete(name="Simulation\nPloidy", breaks=c("sp2","sp4","sp8"), labels=c("2","4","8"))  + facet_grid(cols = vars(traj_ploidy), rows = vars(sampGen))  + xlab("Position (100 kb)") + ylab(expression("-log"[10]*"(p-value)")) + theme_bw()   
```

## Sample generation effect   Does signal decay faster in higher or lower
ploidies?   Answer seems to be that decay is faster in lower ploidies.
Drift is faster in lower ploidies, which is primary mechanism that erases signal
of selection.     
```{r}   
fact.ehh %>% filter(s == "se0.01" &
fuseGen=="fG1" & dom=="do0.5" & N=="NN10000" & recomb=="re1e-08") %>%
ggplot(.,aes(x=POSITION,y=xpehh.p,color=sim_ploidy)) + geom_smooth() +
facet_grid(rows = vars(traj_ploidy), cols = vars(sampGen))   
```

---   # Less interesting analyses   
### Effect of sample
generation   Sample generation is the number of generations following a
sweep at which sampling occurs   
```{r}    
ms2 %>% filter(dom %in%
c("0.1","0.9","0.5") & s==0.01 & fuseGen == 1 & mu==1e-08 & N == 10000 & recomb
== 5e-08 & sim_ploidy == traj_ploidy) %>% ggplot(., aes(y=Pi.1 * 100, color =
as.factor(sim_ploidy), x = mid/100000)) + geom_smooth(method="loess", size =0.5,
span=0.1, se=F) + scale_x_continuous(breaks=c(1,5,9)) +
scale_y_continuous(breaks=c(0.05,0.15)) + xlab("Position (100 kb)") +
ylab("Diversity (%)") + scale_color_discrete(name="Simulation\nPloidy") +
facet_grid(vars(sampGen), vars(dom),labeller = label_both) + theme_bw()  
```

### Effect of population size   Nothing really interesting
here...Smaller population size seems noisier as expected   
```{r}
 ms2 %>% filter(dom %in% c("0.1","0.9","0.5") & recomb=="1e-08" & fuseGen ==
1 & sampGen ==1 & mu==1e-08 & sim_ploidy == traj_ploidy) %>% ggplot(.,
aes(y=Pi.1 * 100, color = as.factor(traj_ploidy), x = mid/100000)) +
geom_smooth(method="loess", size =0.5, span=0.1, se=F) +
scale_x_continuous(breaks=c(1,5,9)) + xlab("Position (100 kb)") +
ylab("Diversity (%)") + scale_color_discrete(name="Ploidy") + facet_grid(vars(s,
N), vars(dom),labeller = label_both, scales = "free_y") + theme_bw()  
``` 
  ### Other plots   
```{r}   fact.ehh %>%
filter(traj_ploidy == "tp4" & fuseGen=="fG1" & dom=="do0.5" & N=="NN10000" &
recomb=="re1e-08" & sampGen=="sG1") %>% ggplot(.,aes(x=cut_interval(x=POSITION,
n=50),y=xpehh.p,fill=sim_ploidy)) + geom_boxplot() + facet_grid(cols =
vars(sampGen))   # # Look for proportional difference in breadth/depth
within a simul   MS2.a %>% filter(fuseGen == 1 & mu==1e-08 & traj_ploidy
== 4 & N == 10000 & recomb=="1e-08" & s =="0.01") %>% ggplot(., aes(y=depth,
fill = as.factor(sampGen), x = as.factor(sim_ploidy))) + geom_boxplot(notch=T) +
ylab(expression(paste("Dip Breadth = Length of which ", pi[1], " < (",pi[2], " /
2)",sep = "")))+ scale_fill_discrete(name = "Trajectory\nPloidy") +
xlab("Simulation Ploidy") + theme_bw()   fact.ehh %>%
filter(sim_ploidy=="sp8" & fuseGen=="fG1" & sampGen=="sG1" & recomb == "re1e-08"
& dom=="do0.5" & N=="NN10000") %>% ggplot(.,aes(x=POSITION /
100000,y=xpehh.p,color=traj_ploidy)) + geom_smooth(method="loess", span=0.1, se
= F) + scale_color_discrete(name="Trajectory\nPloidy",
breaks=c("tp2","tp4","tp8"), labels=c("2","4","8")) + xlab("Position (100 kb)")
+ ylab(expression("-log"[10]*"(p-value)")) + theme_bw()   ```  
```{r}   diffusionDensity = function(start_freq, N, ploidy, t){  
x = seq(0, 1, 0.01)     d = N * ploidy     p = start_freq  
dens = (6 * p * (1 - p) * exp(-t/d)) + (30 * p * (1 - p) * (1 - 2*p) * (1 - 2 *
x ) * exp((-3*t)/d))     return(data.frame("dens" = dens, x = x)) 
 }   aa = diffusionDensity(1/4000, 1000, 4, 1000)   aa %<>%
mutate(ul = x, ll = lag(x), height = (lag(dens) + dens) / 2, width = ul - ll,
area = height * width) %>% filter(area > 0) %>% select(-x) %>% mutate(prob =
area / sum(area))   dfreq = sample(aa$ul, 1, replace = T, prob = aa$prob)
  Drift = function(N, ploidy, end, tries){     drift = function(N,
ploidy, end, tries){       C = N * ploidy # # of chromosomes   p =
0 # Current frequency       traj = c(1/C)       while(p == 0){ 
       p = rbinom(1, C, (1 / C)) / (C) # Draw # of mutant alleles for next
generation and divive by C to get freq       }       t = 1 
     traj = c(traj, p)       while(p != 0 & p != 1 & t < end){ # Loop
till absorption or beginning of selection         p = rbinom(1, C, p) / C
        t = t + 1         traj = c(traj, p)       } 
     return(traj)     }     tt = 1     traj = drift(N,
ploidy, end)     while(tt < tries & (traj[length(traj)] == 0 |
traj[length(traj)] == 1)){       traj = drift(N, ploidy, end)  
}     return(traj)   }   ```
